\scshape Advanced game theory

### 2025-12-07 ch_tar.tex
Add files via upload

\begin{namedthm}[\Cref*{exercise:lb-R2}\ref*{item:lb_R2:a}--\ref*{item:lb_R2:d} {\normalfont (continued)}.]
	%
	\label{exercise:lb-R2_lat_a-d}
	%
	Show that the infimum in $\left( [0,1]^2, \mathord{\leq} \right)$ of any set $S \subseteq [0,1]^2$ exists and is equal to the `coordinate-by-coordinate' infimum, i.e. the vector $\bigl( \inf_{([0,1],\mathord{\leq})} S_1, \inf_{([0,1],\mathord{\leq})} S_2 \bigr)$ where
	%
	\begin{equation*}
		S_1 \coloneqq \left\{ x_1 \in [0,1] : (x_1,x_2) \in S \right\}
		\quad \text{and} \quad
		S_2 \coloneqq \left\{ x_2 \in [0,1] : (x_1,x_2) \in S \right\} .
	\end{equation*}
	%
	Further show that $\left( [0,1]^2, \mathord{\leq} \right)$ is a complete lattice.
	%
\end{namedthm}

@@

It is a general fact that for a product lattice $(X,\mathord{\lesssim}) = \left( \prod_{i \in \mathcal{I}} X_i, \mathord{\lesssim} \right)$, the infimum and supremum in $(X,\mathord{\lesssim})$ of any set $S \subseteq X$, if these exist, are equal to $\bigl( \inf_{(X_i,\mathord{\leq})} S_i \bigr)_{i \in \mathcal{I}}$ and $\bigl( \sup_{(X_i,\mathord{\leq})} S_i \bigr)_{i \in \mathcal{I}}$, where $S_i \coloneqq \left\{ x_i \in X_i : (x_i,x_{-i}) \in S \right\}$.
If $X_i$ is compact for each $i \in \mathcal{I}$, then by \Cref{exercise:lb-R_sublatt}\ref{item:lb_R_sublatt:subcompl}, we may replace the `$\inf_{(X_i,\mathord{\leq})}$' and `$\sup_{(X_i,\mathord{\leq})}$' with the ordinary `$\inf$' and `$\sup$' from real analysis.

@@

\begin{namedthm}[\Cref*{exercise:arrow-pratt}\ref*{item:arrow-pratt:b} {\normalfont (continued from \cpageref{exercise:arrow-pratt,item:arrow-pratt:b})}.]
	%
	\label{exercise:arrow-pratt_lattice}
	%
	Show that $\left(\mathcal{X},\text{`less risk-seeking than'}\right)$ is a lattice.
	%
\end{namedthm}

@@

	Let $\mathcal{P}$ be the set of all product subsets of $A$. The map $\mathcal{B} : \mathcal{P} \to \mathcal{P}$ is evidently $\subseteq$/$\subseteq$-increasing. It is easily checked (echoing \Cref{example:meet_boolean} on \cpageref{example:meet_boolean}) that $(\mathcal{P},\mathord{\subseteq})$ is a complete lattice. Hence by \hyperref[theorem:tarski]{Tarski's fixed-point theorem}, there is a largest (in the set-inclusion sense) set with the best-reply property. This recovers \Cref{observation:br_property_max} (\cpageref{observation:br_property_max}). Note that we made no assumptions about the game $(I,(A_i,u_i)_{i \in I})$, such as finiteness or compactness-and-continuity.
	%
\end{example}

>>>>

\begin{namedthm}[\Cref*{exercise:lb-R2}\ref*{item:lb_R2:a}--\ref*{item:lb_R2:d} {\normalfont (continued)}.]
	%
	\label{exercise:lb-R2_lat_a-d}
	%
	Show that the infimum in $\left( [0,1]^2, \mathord{\leq} \right)$ of any set $S \subseteq [0,1]^2$ exists and is equal to the `coordinate-by-coordinate' infimum, i.e. the vector $\bigl( \inf_{([0,1],\mathord{\leq})} S_1, \inf_{([0,1],\mathord{\leq})} S_2 \bigr)$ where
	%
	\begin{align*}
		S_1 &\coloneqq \left\{ x_1 \in [0,1] : \text{$(x_1,x_2) \in S$ for some $x_2 \in [0,1]$} \right\}
		\quad \text{and}
		\\
		S_2 &\coloneqq \left\{ x_2 \in [0,1] : \text{$(x_1,x_2) \in S$ for some $x_1 \in [0,1]$} \right\} .
	\end{align*}
	%
	Further show that $\left( [0,1]^2, \mathord{\leq} \right)$ is a complete lattice.
	%
\end{namedthm}

@@

It is a general fact that for a product lattice $(X,\mathord{\lesssim}) = \left( \prod_{i \in \mathcal{I}} X_i, \mathord{\lesssim} \right)$, the infimum and supremum in $(X,\mathord{\lesssim})$ of any set $S \subseteq X$, if these exist, are equal to $\bigl( \inf_{(X_i,\mathord{\leq})} S_i \bigr)_{i \in \mathcal{I}}$ and $\bigl( \sup_{(X_i,\mathord{\leq})} S_i \bigr)_{i \in \mathcal{I}}$, where
%
\begin{equation*}
	S_i \coloneqq \left\{ x_i \in X_i : \text{$(x_i,x_{-i}) \in S$ for some $x_{-i} \in \prod_{j \in \mathcal{I} \setminus \{i\}} X_j$} \right\} .
\end{equation*}
%
If $X_i$ is compact for each $i \in \mathcal{I}$, then by \Cref{exercise:lb-R_sublatt}\ref{item:lb_R_sublatt:subcompl}, we may replace the `$\inf_{(X_i,\mathord{\leq})}$' and `$\sup_{(X_i,\mathord{\leq})}$' with the ordinary `$\inf$' and `$\sup$' from real analysis.

@@

\begin{namedthm}[\Cref*{exercise:arrow-pratt}\ref*{item:arrow-pratt:b} {\normalfont (continued from \cpageref{exercise:arrow-pratt,item:arrow-pratt:b})}.]
	%
	\label{exercise:arrow-pratt_lattice}
	%
	Show that the poset $\left(\mathcal{X},\text{`less risk-seeking than'}\right)$ is a lattice.
	%
\end{namedthm}

@@

	Let $\mathcal{Y}$ be the set of all product subsets of $A$. The map $\mathcal{B} : \mathcal{Y} \to \mathcal{Y}$ is evidently $\subseteq$/$\subseteq$-increasing. It is easily checked (echoing \Cref{example:meet_boolean} on \cpageref{example:meet_boolean}) that $(\mathcal{Y},\mathord{\subseteq})$ is a complete lattice. Hence by \hyperref[theorem:tarski]{Tarski's fixed-point theorem}, there is a largest (in the set-inclusion sense) fixed point of $\mathcal{B}$. The largest fixed point $\mathcal{B}$ is also the largest set with the best-reply property (why?). This recovers \Cref{observation:br_property_max} (\cpageref{observation:br_property_max}). Note that we made no assumptions about the game $(I,(A_i,u_i)_{i \in I})$, such as finiteness or compactness-and-continuity.
	%
\end{example}

### 2025-12-07 ms.tex
Add files via upload

% watermark on every page (for version control)
% \usepackage[anchor=ll,pos={0.1cm,0.5cm},fontsize=0.2cm,angle=0,alignment=l]{draftwatermark}
% \SetWatermarkText{\normalfont\hspace{0.37em}version:\\\normalfont {\datestyle\today}\\\normalfont\hspace{0.37em}at {\currenttime}}

@@

\date{\small This version: 16 April 2025}

@@

\makeatletter
	\AtBeginDocument{ \hypersetup{
		pdftitle = {Game theory},
		pdfauthor = {Ludvig Sinander}
		} }
\makeatother

@@

An \emph{extensive-form game} is a tuple $(I,\mathcal{A},H,P,A,\mathcal{S},(u_i)_{i \in I},\pi)$, where

@@

	\item $H$ is a set of sequences in $\mathcal{A}$ that contains the empty sequence $\varnothing$ and is closed under truncation (that is, if $h$ is a truncation of $h' \in H$ then $h \in H$), where we write $Z$ for the set of terminal elements of $H$,

@@

	\item $A$ is a map $H \setminus Z \to 2^\mathcal{A} \setminus \{\varnothing\}$ such that $A(H \setminus Z) = \mathcal{A}$,

@@

	\item for each $h \in H \setminus Z$ such that $P(h) = \text{Nature}$, $\pi(h)$ is a full-support probability distribution over $A(h)$ (that is, an element of $\Delta(A(h))$).%
		\footnote{Thus, explicitly, $\pi$ is a map $P^{-1}(\text{Nature}) \to \Union_{h \in P^{-1}(\text{Nature})} \Delta(A(h))$ such that for each $h \in P^{-1}(\text{Nature})$ and $a \in \Union_{h \in P^{-1}(\text{Nature})} A(h)$, $\pi(h)(a)>0$ iff $a \in A(h)$, where $P^{-1}(\text{Nature}) \equiv \{ h \in H \setminus Z : P(h) = \text{Nature} \}$.}

\end{itemize}
%
The interpretation is that $I$ is the set of players, $\mathcal{A}$ is the set of all actions, $H$ is the set of histories (possible action sequences), and the functions $P$ and $A$ specify which player moves at each history and what actions are then available. The elements of the partition $\mathcal{S}$ are called \emph{information sets.} These describe what players observe: the player $P(h)$ who moves at history $h \in H \setminus Z$ observes which information set $S \in \mathcal{S}$ the history $h$ belongs to, but does not observe any more than that. The functions $(u_i)_{i \in I}$ are agents' payoff functions, defined on terminal histories (meaning that payoffs can in general depend on the entire history of play). Finally, it may be that at some histories $h \in H \setminus Z$, the move is made by $P(h) = \text{Nature}$, who chooses an action from $A(h)$ according to an exogenous commonly-known full-support probability distribution $\pi(h) \in \Delta(A(h))$. (`Nature' is not a strategic player, just a device for modelling chance moves.)

@@

A \emph{subgame} is the game obtained by reaching some history $h \in H \setminus Z$ that belongs to a singleton information set ($S = \{h\}$ for some $S \in \mathcal{S}$) and looking forward. Formally, a subgame of $(I,\mathcal{A},H,P,A,\mathcal{S},(u_i)_{i \in I},\pi)$ is an extensive-form game $(I',\mathcal{A}',H',P',A',\mathcal{S}',(u_i')_{i \in I},\pi')$ such that for some $h_0 \in H \setminus Z$,
%
\begin{itemize}

	\item $H' = \left\{ h' \in H : \text{$h_0$ is a truncation of $h'$} \right\}$, and we write $Z'$ for the set of terminal elements of $H'$,

@@

	\item $A'$ equals the restriction of $A$ to $H' \setminus Z'$,

>>>>

% % watermark on every page (for version control)
% \usepackage[anchor=ll,pos={0.1cm,0.5cm},fontsize=0.2cm,angle=0,alignment=l]{draftwatermark}
% \SetWatermarkText{\normalfont\hspace{0.37em}version:\\\normalfont {\datestyle\today}\\\normalfont\hspace{0.37em}at {\currenttime}}

@@

\date{\small This version: 7 December 2025}

@@

\makeatletter
	\AtBeginDocument{ \hypersetup{
		pdftitle = {Advanced game theory},
		pdfauthor = {Ludvig Sinander}
		} }
\makeatother

@@

An \emph{extensive-form game} is a tuple $(I,\mathcal{A},H,P,\mathcal{S},(u_i)_{i \in I},\pi)$, where

@@

	\item $H$ is a set of sequences in $\mathcal{A}$ that contains the empty sequence $\varnothing$, is closed under truncation (that is, if $h$ is a truncation of $h' \in H$ then $h \in H$), and features every element of $\mathcal{A}$ (that is, for each $a \in \mathcal{A}$, there exists a history $h = (a_n)_{n=1}^N \in H$ such that $a_n = a$ for some $n$), where we write $Z$ for the set of terminal elements of $H$ and $A(h) \coloneqq \left\{ a \in \mathcal{A} : (h,a) \right\}$ for the set of actions available at a non-terminal history $h \in H \setminus Z$,

@@

@@

	\item for each $h \in H \setminus Z$ such that $P(h) = \text{Nature}$, $\pi(h)$ is a full-support probability distribution over $A(h)$ (that is, an element of the interior of $\Delta(A(h))$).%
		\footnote{Thus, explicitly, $\pi$ is a map $P^{-1}(\text{Nature}) \to \Union_{h \in P^{-1}(\text{Nature})} \Delta(A(h))$ such that for each $h \in P^{-1}(\text{Nature})$ and $a \in \Union_{h \in P^{-1}(\text{Nature})} A(h)$, $\pi(h)(a)>0$ iff $a \in A(h)$, where $P^{-1}(\text{Nature}) \equiv \{ h \in H \setminus Z : P(h) = \text{Nature} \}$.}

\end{itemize}
%
The interpretation is that $I$ is the set of players, $\mathcal{A}$ is the set of all actions, $H$ is the set of histories (possible action sequences), and the function $P$ specifies which player moves at each history. The elements of the partition $\mathcal{S}$ are called \emph{information sets.} These describe what players observe: the player $P(h)$ who moves at history $h \in H \setminus Z$ observes which information set $S \in \mathcal{S}$ the history $h$ belongs to, but does not observe any more than that. The functions $(u_i)_{i \in I}$ are agents' payoff functions, defined on terminal histories (meaning that payoffs can in general depend on the entire history of play). Finally, it may be that at some histories $h \in H \setminus Z$, the move is made by $P(h) = \text{Nature}$, who chooses an action from $A(h)$ according to an exogenous commonly-known probability distribution $\pi(h) \in \Delta(A(h))$ with full support (on $A(h)$). (`Nature' is not a strategic player, just a device for modelling chance moves.)

@@

A \emph{subgame} is the game obtained by reaching some history $h \in H \setminus Z$ that belongs to a singleton information set ($S = \{h\}$ for some $S \in \mathcal{S}$) and looking forward. Formally, a subgame of $(I,\mathcal{A},H,P,\mathcal{S},(u_i)_{i \in I},\pi)$ is an extensive-form game $(I',\mathcal{A}',H',P',\mathcal{S}',(u_i')_{i \in I},\pi')$ such that for some $h_0 \in H \setminus Z$,
%
\begin{itemize}

	\item $H' = \left\{ h' \in H : \text{$h_0$ is a truncation of $h'$} \right\}$, and we write $Z'$ for the set of terminal elements of $H'$ and $A'$ for the restriction of $A$ to $H' \setminus Z'$,

@@
